---
title: "Supporting Material A"
author: "Shinichi Nakagawa, Malgorzata Lagisz, Rose E. O’Dea, Patrice Pottier, Joanna Rutkowska, Alistair M. Senior, Yefeng Yang, and Daniel W.A. Noble"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_download: true
    code_folding: hide
    depth: 4
    number_sections: no
    theme:  cosmo # “default”, “cerulean”, “journal”, “flatly”, “darkly”, “readable”, “spacelab”, “united”, “cosmo”, “lumen”, “paper”, “sandstone”, “simplex”, “yeti”
    toc: yes
    toc_float: yes
    toc_depth: 4
  pdf_document:
    toc: yes
subtitle:  "orchaRd 2.0: An R package for visualizing meta-analyses with orchard plots"
#bibliography: references.bib
#biblio-style: "apalike"
#csl: nature.csl
link-citations: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, eval = TRUE}
#knitter setting
knitr::opts_chunk$set(
message = FALSE,
warning = FALSE, # no warnings
cache = TRUE,# Caching to save time when knitting
tidy = TRUE, 
echo = TRUE
#fig.width = 9
)

# cleaning up
#rm(list=ls())
```

## Setup
Loading packages and custom functions.   

```{r load packages, eval = TRUE}
pacman::p_load(tidyverse, # tidy family and related packages below
               readxl, #read Excel files
               kableExtra, # add features to a kable() output 
               purrr,    #functional programming tools
               magrittr, # extending piping
               pander,   # nice tables
               plotly,     # interactive plots using ggplot2
               formatR, # provides a function tidy_source() to format R source automatically
               here #simplifying file paths
)
```


# Supplementary Methods and Results - Literature survey 

 
## Literature search - methods   

The original literature survey was conducted in 2019 to examine how forest plots are used in ecology and evolution, and what features they have. The resulting dataset was used in the original orchard plot publication:   

Nakagawa, S., Lagisz, M., O'Dea, R. E., Rutkowska, J., Yang, Y., Noble, D. W., & Senior, A. M. (2020). The Orchard Plot: Cultivating a Forest Plot for Use in Ecology, Evolution and Beyond. Research Synthesis Methods https://doi.org/10.1002/jrsm.1424 12: 4-12 (EcoEvoRxiv https://doi.org/10.32942/osf.io/epqa7)  

Here is a brief description of the systematic search of ecology and evolution journals underlying the data set from the SI materials from the above paper: "In brief, we collected the ISSN of journals classified under ‘Ecology’ and/or ‘Evolutionary Biology’, according to the ISI InCites Journal Citation Reports. We then searched within these journals for papers published between 1 January 2010 and 25 March 2019, and which contained any of the following search terms in the title, abstract, or keywords: “meta-analy* “, “metaanaly* “, or “meta-regression”. To reduce the large pool of papers to a representative sample, we filtered the results to only include the subset of journals which often published meta-analyses: the top 10 most represented ecology journals; the top 10 most represented evolution journals, and; the top 11 most represented journals that were classified as both ecology and evolutionary biology. We then took a random sample of papers from each journal to give 297 papers for screening. During screening we identified 102 papers that unambiguously addressed a question in the fields of ecology or evolutionary biology, performed a search for and collected data from primary literature, extracted or calculated the same type of effect size from multiple studies, and analysed those effect sizes in a statistical model. These criteria identify a representative sample of 102 ‘traditional’ published meta-analyses (Extended Data Table 1).   

**Extended Data Table 1:**   
List of 102 representative meta-analytic papers published in ecology and/or evolution journals.      

```{r list of papers, include = TRUE, eval = TRUE}
# getting the papers data and formatting some variables (turning character vectors to factors)
papers_data <- read_excel(here("data", "forest_plot_fulltext_records.xlsx"), na = "NA") %>% 
   mutate_if(is.character, as.factor)
#str(papers_data) #102 papers
#retain fulltext_ID, authors, pub_year, title, journal, volume, pages and make scrollable table:
papers_data %>% select(fulltext_ID, authors, pub_year, title, journal, volume, pages) %>% kable("html") %>%
    kable_styling("striped", position = "left") %>%
  scroll_box(width = "100%", height = "300px")
```
" (end of quotation)

 

## Literature survey for orchard plot pacakge update      

We designed a new data survey to answer the following questions relevant to the new features of the orchard plot package:   
   
 1.	How many papers (percentage of 102 surveyed) have at least one categorical variable/moderator? [This could be a variable/moderator that was used in analyses or to subset the data for analyses.]
 
 2.  How many papers (percentage of 102 surveyed) have at least one test or model for heteroskedasticity (heteroscedasticity)? [i.e. test for homogeneity of variance or heterogeneous variance between different groups.] 
 
 3. How many papers (percentage of 102 surveyed) have at least one model in a paper with more than one categorical moderator? [Note, this should be a multi-moderator model with at least two categorical moderators.]   
 
 4. How many papers (percentage of 102 surveyed) have at least one model with at least one categorical moderator and one continuous moderator? [Note, this is only for predictors variables other than SE (sample size) or publication year.] 
 
 5. How many papers (percentage of 102 surveyed) that used a multi-moderator regression have at least one plot (figure) made from a multi-moderator meta-regression? [This is a plot where plotted values are marginalized other variables in a multi-moderator meta-regression model]. So many people will make figures from uni-moderator meta-regression. 
 
 6. How many papers (percentage of 102 surveyed) that used a multi-moderator regression also modeled interactions?    
 
 7.  How many papers, which used R, cite an R software package they used for meta-analysis?   
 
 8.  How many papers, which used R, cite an R software package they used for graphical presentation of meta-analytic results?  


### Data extraction
We designed a custom data extraction form to gather information on the types of models and plots used in 102 included articles and implemented it as a Google Form.  We based our classification on the information present in the full texts and supplementary files of the articles.     

Two researchers (from a pool of three: ML, JR, YY) independently extracted data from all articles. All three reviewers discussed and resolved any disagreements in the coded data. Consensus data was used for the analyses.   

## Literature survey - Results   
Data loading and preprocessing. Prepocessed data is presented in Extended Data Table 2.   

```{r load coded data, eval = TRUE}
# getting the data and formatting some variables (turning character vectors to factors)
full_data <- read_csv(here("data", "orchaRd2 (Responses) - Consesus Responses 2.csv"), na = "NA") %>%    mutate_if(is.character, as.factor)
#str(full_data) #each row represents one scoring attempt, so it is usually 2 rows per figure file (since 2 people were assessing independently)
```

```{r clean coded data, include = FALSE, eval = TRUE}
#names(full_data)
#View(full_data)

#create a copy to be reduced to data used for analyses, also remove unnecessary columns and rename the other
dat <- full_data  %>% select(-starts_with("Comment"), -"Timestamp", -"Respondents initials")
dat <- dat  %>% rename(Article_id = "Article id", 
                       One_categorical_var = Q1, 
                       Test_heteroskedasticity = Q2, 
                       Many_categorical_var = Q3, 
                       Categorical_cont_var = Q4, 
                       Multimoderator_plot = Q5,
                       Interactions_mod = Q6, 
                       Cited_ma_package = Q7, 
                       Cited_plot_package = Q8
                       )
#names(dat)

#check Article_id
length(unique(dat$Article_id)) #102
all <- unique(dat$Article_id) #store IDs of all papers

# #recode factor levels into shorter text, where needed:
# levels(dat$Metaregression_subgroup) #ok
# levels(dat$One_categorical_var) #ok
# levels(dat$Many_categorical_var) #ok
# levels(dat$Test_heteroskedasticity) #ok
# levels(dat$Many_categorical_var)
# levels(dat$Categorical_cont_var) #ok
# levels(dat$Interactions_mod) #this one will require splitting (multiple values)
# levels(dat$Multimoderator_plot) #ok
# levels(dat$Cited_ma_package) #ok
# levels(dat$Cited_plot_package) #ok
# levels(dat$Cited_extr_package) #ok

#re-arrange columns
dat <- dat %>%
  select(Article_id,One_categorical_var, Test_heteroskedasticity, Many_categorical_var, Categorical_cont_var, Multimoderator_plot, Interactions_mod, Cited_ma_package, Cited_plot_package)
```


**Extended Data Table 2:** 
Preprocessed coded data from the literature survey of 102 meta-analytic papers in ecology or evolution.

```{r table of coded data, eval = TRUE}
# making a scrollable table
kable(dat, "html") %>%
  kable_styling("striped", position = "left") %>%
  scroll_box(width = "100%", height = "300px")
```

### 1.	How many papers have at least one categorical variable/moderator?   
   
**Answer 1.**	   
Almost all `r sum(dat$One_categorical_var == "Yes")` papers (`r round(sum(dat$One_categorical_var == "Yes")/102*100, 2)`%, out of 102 surveyed papers) had at least one categorical variable/moderator. [Working definition used: categorical variable/moderator could be a variable/moderator that was used in analyses or to subset the data for analyses.]     

```{r summarise One_categorical_var, eval = TRUE, fig.height = 2}
#table(dat$One_categorical_var, useNA = "always")
dat %>% 
  group_by(One_categorical_var) %>% 
  count() %>%
  ggplot(aes(x = One_categorical_var, y = n)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(margin = margin(0, 5, 0, 60))) + 
  xlab("") +
  ylab("Number of articles") +
  ylim(0, 100)
```


### 2.	How many papers have at least one test for heteroskedasticity (heteroscedasticity)?   
   
**Answer 2.**	   
Only `r sum(dat$Test_heteroskedasticity == "Yes")` papers (`r round(sum(dat$Test_heteroskedasticity == "Yes", na.rm = TRUE)/102*100, 2)`%, out of 102 surveyed papers) explicitly tested for heteroskedasticity in their data. [[i.e. tested for homogeneity of variance or heterogeneous variance between different groups.]     

```{r summarise Test_heteroskedasticity, eval = TRUE, fig.height = 2}
#table(dat$Test_heteroskedasticity, useNA = "always")
dat %>% 
  group_by(Test_heteroskedasticity) %>% 
  count() %>%
  ggplot(aes(x = Test_heteroskedasticity, y = n)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(margin = margin(0, 5, 0, 60))) + 
  xlab("") +
  ylab("Number of articles") +
  ylim(0, 100)
```



### 3.	How many papers have at least one model in a paper with more than one categorical moderator?    
   
**Answer 3.**	   
Almost half of the surveyed papers (`r round(sum(dat$Many_categorical_var == "Yes")/102*100, 2)`%, out of 102 surveyed papers) had more than one categorical variable/moderator included in a statistical model.   

```{r summarise Many_categorical_var, eval = TRUE, fig.height = 2}
#table(dat$Many_categorical_var, useNA = "always")
dat %>% 
  group_by(Many_categorical_var) %>% 
  count() %>%
  ggplot(aes(x = Many_categorical_var, y = n)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(margin = margin(0, 5, 0, 60))) + 
  xlab("") +
  ylab("Number of articles") +
  ylim(0, 100)
```


### 4.	How many papers have at least one model with at least one categorical moderator and one continuous moderator?  
   
**Answer 4.**	   
Around quarter (`r round(sum(dat$Categorical_cont_var == "Yes", na.rm = TRUE)/102*100, 2)`%, out of 102 surveyed papers) of papers had at least one categorical variable/moderator. [Note, this is only for predictors variables other than SE (sample size) or publication year.]     

```{r summarise Categorical_cont_var, eval = TRUE, fig.height = 3}
#table(dat$Categorical_cont_var, useNA = "always")
dat %>% 
  group_by(Categorical_cont_var) %>% 
  count() %>%
  ggplot(aes(x = Categorical_cont_var, y = n)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(margin = margin(0, 5, 0, 60))) + 
  xlab("") +
  ylab("Number of articles") +
  ylim(0, 100)
```


### 5.	How many papers that used a multi-moderator regression have at least one plot (figure) made from a multi-moderator meta-regression?     
   
**Answer 5.**	   
Majority of the included papers (`r round(sum(dat$Multimoderator_plot == "No", na.rm = TRUE)/102*100, 2)`%, out of 102 surveyed papers) did not visualise marginalised results of a multi-moderator regression in a plot (figure).   

```{r summarise Multimoderator_plot, eval = TRUE, fig.height = 3}
#table(dat$Multimoderator_plot, useNA = "always")
dat %>% 
  group_by(Multimoderator_plot) %>% 
  count() %>%
  ggplot(aes(x = Multimoderator_plot, y = n)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(margin = margin(0, 5, 0, 60))) + 
  xlab("") +
  ylab("Number of articles") +
  ylim(0, 100)
```


### 6.	How many papers that used a multi-moderator regression also modeled interactions?    
   
**Answer 6.**	   
Most (`r sum(dat$Interactions_mod == "No", na.rm = TRUE) + sum(is.na(dat$Interactions_mod))`) papers (`r round((sum(dat$Interactions_mod == "No", na.rm = TRUE) + sum(is.na(dat$Interactions_mod)))/102*100, 2)`%, out of 102 surveyed papers) did not model interactions as part of their statistical models. When interactions were included (one paper could have tested more than one type of interactions), continuous x continuous variable interactions were least frequent (`r round(7/30*100, 2)`%, out of 30  papers with interactions) . [Note: by interactions we mean statistical interactions between variables in a model, not biological/ecological interactions analysed as effect sizes, e.g. between species.]     

```{r summarise Interactions_mod, eval = TRUE, fig.height = 3}
#table(dat$Interactions_mod, useNA = "always")
dat$Interactions_mod <- str_replace_all(dat$Interactions_mod, ", ", "; ") #fix  separators

dat %>% 
  filter(!is.na(Interactions_mod)) %>%
  filter(Interactions_mod != "No") %>% 
  pull(Interactions_mod) %>% 
  strsplit("; ") %>% 
  unlist() -> Interactions_mod_sep

## fix typos and inconsistencies:
Interactions_mod_sep <- gsub("continious x categorical", "categorical x continious", Interactions_mod_sep)
Interactions_mod_sep <- gsub("continious x continuous", "continious x continious", Interactions_mod_sep)

table(Interactions_mod_sep) %>% as.data.frame() %>%
  ggplot(aes(x = Interactions_mod_sep, y = Freq)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(margin = margin(0, 5, 0, 30))) + 
  scale_x_discrete(labels = c("categorical x\n categorical","categorical x\n continous","continous x\n continous"))+ 
  xlab("") +
  ylab("Number of articles") +
  ylim(0, 100)
```

### 7.	How many papers cite an R software package they used for meta-analysis?      
   
**Answer 7.**	   
R packages used for meta-analysis were commonly cited (`r sum(dat$Cited_ma_package == "Yes")` papers, `r round(sum(dat$Cited_ma_package == "Yes")/((sum(dat$Cited_ma_package == "Yes")+sum(dat$Cited_ma_package == "No")))*100, 2)`%, out of `r ((sum(dat$Cited_ma_package == "Yes")+sum(dat$Cited_ma_package == "No")))`` surveyed papers that used R).      

```{r summarise Cited_ma_package, eval = TRUE, fig.height = 4}
#table(dat$Cited_ma_package, useNA = "always")

## set the levels in order we want:
dat$Cited_ma_package <- factor(dat$Cited_ma_package, levels = rev(c("Yes", "No", "Unclear", "Did not use R")))

dat %>% 
  group_by(Cited_ma_package) %>% 
  count() %>%
  ggplot(aes(x = Cited_ma_package, y = n)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(margin = margin(0, 5, 0, 30))) + 
  xlab("") +
  ylab("Number of articles") +
  ylim(0, 100)
```


### 8.	How many papers cite an R software package they used for graphical presentation of meta-analytic results?  
   
**Answer 8.**	   
Only `r sum(dat$Cited_plot_package == "Yes")` papers (`r round(sum(dat$Cited_plot_package == "Yes")/(sum(dat$Cited_ma_package == "Yes")+sum(dat$Cited_ma_package == "No"))*100, 2)`%, out of `r sum(dat$Cited_ma_package == "Yes")+sum(dat$Cited_ma_package == "No")` surveyed papers that used R) cited R package used to present meta-analytic results.    

```{r summarise Cited_plot_package, eval = TRUE, fig.height = 4}
#table(dat$Cited_plot_package, useNA = "always")

## set the levels in order we want:
dat$Cited_plot_package <- factor(dat$Cited_plot_package, levels = rev(c("Yes", "No", "Unclear", "Did not use R")))

dat %>% 
  group_by(Cited_plot_package) %>% 
  count() %>%
  ggplot(aes(x = Cited_plot_package, y = n)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(margin = margin(0, 5, 0, 40))) + 
  xlab("") +
  ylab("Number of articles") +
  ylim(0, 100)
```

